# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r get_dataset, echo = TRUE}
dataset = read.csv("activity.csv")
dataset$date = strptime(dataset$date, "%Y-%m-%d")
head(dataset)
```

## What is mean total number of steps taken per day?
clean.dataset: ignore missing values
```{r}
clean.dataset = na.omit(dataset) 
head(clean.dataset)
```

Histogram of the total number of steps taken each day
```{r}
library(plyr)
steps.date = ddply(clean.dataset, ~date, summarize, sum=sum(steps))
colnames(steps.date) = c("date", "steps")
hist(steps.date$steps)
```

The **mean** and **median** total number of steps taken per day
```{r}
steps.date.mean = ddply(clean.dataset, ~date, summarize, 
                        mean = mean(steps),
                        median = median(steps))
steps.date.mean
```

## What is the average daily activity pattern?
A time series plot of the 5-minute interval (x-axis) and 
the average number of steps taken, averaged across all days (y-axis)
```{r}
steps.interval.mean = ddply(clean.dataset, ~interval, summarize, 
                        mean = mean(steps))
library(ggplot2)
g = ggplot(steps.interval.mean, aes(interval, mean))
g + geom_line()
```

The 5-minute interval, on average across all the days 
in the dataset, that contains the maximum number of steps:
```{r}
steps.interval.mean[steps.interval.mean$mean ==
                      max(steps.interval.mean$mean), "interval"]
```

## Imputing missing values
The total number of rows with missing values in the dataset:
```{r}
nrow(dataset[is.na(dataset), ])
```

Going to impute each NA steps with mean of that 5-min interval across the rest of  dataset.

impute.dataset: new dataset that is equal to the original dataset 
but with the missing data filled in.
```{r}
impute.dataset = merge(dataset, steps.interval.mean, by="interval")
for (i in 1:nrow(impute.dataset)) {
  if(is.na(impute.dataset[i,"steps"])) {
    impute.dataset[i,"steps"] = impute.dataset[i,"mean"]
  }
}
impute.dataset$mean=NULL
head(impute.dataset)
```

Compare with dataset where missing values were ignored.
impute.dataset, Histogram of the total number of steps taken each day
```{r}
steps.date1 = ddply(impute.dataset, ~date, summarize, sum=sum(steps))
colnames(steps.date1) = c("date", "steps")
hist(steps.date1$steps)
```

impute.dataset, The **mean** and **median** total number of steps taken per day
```{r}
steps.date.mean1 = ddply(impute.dataset, ~date, summarize, 
                        mean = mean(steps),
                        median = median(steps))
steps.date.mean1
```

Comparison of mean and median of both datasets
```{r}
merge(steps.date.mean, steps.date.mean1, by="date", all.y=TRUE)
```
All days have mean and median now.


## Are there differences in activity patterns between weekdays and weekends?

day.type: New factor variable "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
impute.dataset$weekdays = weekdays(impute.dataset$date)
impute.dataset$day.type = "weekday"
impute.dataset[impute.dataset$weekdays == "Saturday" | 
                 impute.dataset$weekdays == "Sunday",
               "day.type"] = "weekend"
impute.dataset$day.type = as.factor(impute.dataset$day.type)
head(impute.dataset)
impute.dataset = impute.dataset[order(impute.dataset$date, impute.dataset$interval),]
head(impute.dataset)
```

Plot of weekdays and weekends
```{r}
steps.interval.mean1 = ddply(impute.dataset, .(interval,day.type), summarize, 
                            mean = mean(steps))

g = ggplot(steps.interval.mean1, aes(interval, mean))
g + geom_line() + facet_wrap(~ day.type, nrow = 2, ncol = 1) 
```
Yes, there is difference in activity between weekdays and weekends. the max of weekday is higher but it seems the weekend is more active overall.
